Between-dataset comparisons

### 2019-11-14 analysis/reports/cross_dataset.Rmd
cross dataset report stuff

```{r overall percentile heatmap}

overall_heatmap_skew <- ggplot(data = all_di, aes(x = s0, y = n0, color = skew_percentile)) +
  geom_point(alpha = .2) +
  theme_bw() +
  scale_color_viridis_c(option = "plasma", begin = .1, end = .8, direction = -1) +
  facet_wrap(vars(singletons)) +
  ggtitle("Skewness Percentile") +
  theme(legend.position = "bottom")

overall_heatmap_simpson <- ggplot(data = all_di, aes(x = s0, y = n0, color = simpson_percentile)) +
  geom_point(alpha = .2) +
  theme_bw() +
  scale_color_viridis_c(option = "plasma", begin = .1, end = .8)+
  facet_wrap(vars(singletons)) +
  ggtitle("Simpson's Percentile")+
  theme(legend.position = "bottom")

overall_heatmap_skew_zoomed <- overall_heatmap_skew +
  ylim(0, 5000) +
  xlim(0, 250)+
  theme(legend.position = "none")

overall_heatmap_simpson_zoomed <- overall_heatmap_simpson +
  ylim(0, 5000) +
  xlim(0, 250)+
  theme(legend.position = "none")

gridExtra::grid.arrange(grobs = list(overall_heatmap_skew, overall_heatmap_skew_zoomed), nrow= 1)
gridExtra::grid.arrange(grobs = list(overall_heatmap_simpson, overall_heatmap_simpson_zoomed), nrow= 1)

Broken out by dataset

```{r dataset heatmaps, fig.width = 30, fig.height = 20}

dat_heatmap_skew <-  ggplot(data = filter(all_di, !singletons), aes(x = s0, y = n0, color = skew_percentile)) +
  geom_point(alpha = .4) +
  theme_bw() +
  scale_color_viridis_c(option = "plasma", begin = .1, end = .8, direction = -1)+
  facet_wrap(vars(dat), scales = "free") +
  ggtitle("Skew Percentile")+
  theme(legend.position = "bottom")
dat_heatmap_skew

dat_heatmap_simpson <-  ggplot(data = filter(all_di, !singletons), aes(x = s0, y = n0, color = simpson_percentile)) +
  geom_point(alpha = .4) +
  theme_bw() +
  scale_color_viridis_c(option = "plasma", begin = .1, end = .8)+
  facet_wrap(vars(dat), scales = "free") +
  ggtitle("Simpson's Percentile")+
  theme(legend.position = "bottom")
dat_heatmap_simpson
```

Against S0

```{r s0 percentile}

overall_s0_skew <- ggplot(data = filter(all_di, !singletons), aes(x = s0, y = skew_percentile)) +
  geom_point(alpha = .1) +
  geom_smooth(method = "lm", se = F, color = "pink") +
  theme_bw() +
  ggtitle("Skew percentile")
overall_s0_skew

dat_s0_skew <- overall_s0_skew +
  facet_wrap(vars(dat), scales = "free")
dat_s0_skew

overall_s0_simpson <- ggplot(data = filter(all_di, !singletons), aes(x = s0, y = simpson_percentile)) +
  geom_point(alpha = .1) +
  geom_smooth(method = "lm", se = F, color = "pink") +
  theme_bw() +
  ggtitle("Simpson percentile")
overall_s0_simpson

dat_s0_simpson <- overall_s0_simpson +
  facet_wrap(vars(dat), scales = "free")
dat_s0_simpson

Against nsamples

```{r nsamples}

skew_nsamples <- ggplot(data = filter(all_di, !singletons), aes(x = nsamples, y = skew_percentile)) +
  geom_point(alpha = .1) +
    geom_smooth(method = "lm", se = F, color = "pink") +

  theme_bw() +
  ggtitle("Skew percentile v. nsamples")

skew_nsamples

dat_skew_nsamples <- skew_nsamples +
  facet_wrap(vars(dat))

dat_skew_nsamples

simp_nsamples <- ggplot(data = filter(all_di, !singletons), aes(x = nsamples, y = simpson_percentile)) +
  geom_point(alpha = .1) + 
    geom_smooth(method = "lm", se = F, color = "pink") +

  theme_bw() +
  ggtitle("Simpson percentile v. nsamples")

simp_nsamples
```

@@

>>>>

@@

Observations from these histograms:

* FIA really struggles to get 2500 samples
* Portal is the *most* concentrated of all the datasets.
* Then MCDB and misc, then gentry, then bbs. FIA is hard to place but looks idiosyncratic.
* Curious about how variation in %ile maps on to S and N, *particularly* for Gentry and BBS. 
* Skewness is less extreme than Simpson (more density in the in-between). 
* Singletons doesn't appear to be terrifically impactful. From now on filtering to *unaltered* vectors.

```{r s and n mapping, fig.width = 12, fig.height = 10}

all_di <- filter(all_di, !singletons)

skew_s_n_maps <- ggplot(data = all_di, aes(x = s0, y = n0, color = skew_percentile)) +
  geom_point(alpha = .3) +
  theme_bw() +
  facet_wrap(vars(dat), scales = "free") + 
  ggtitle("Skewness") +
  scale_color_viridis_c(option = "plasma", end= .75)

skew_s_n_maps

simpson_s_n_maps <- ggplot(data = all_di, aes(x = s0, y = n0, color = simpson_percentile)) +
  geom_point(alpha = .3) +
  theme_bw() +
  facet_wrap(vars(dat), scales = "free") + 
  ggtitle("Simpson") +
  scale_color_viridis_c(option = "plasma", end= .75, direction = -1)

simpson_s_n_maps

The less extreme (low skew, high simpson) vectors appear vaguely collected in the lower right for Gentry and BBS. Those are the regions with relatively high N/S, aka low average abundance, aka a relatively small feasible set.

Conversely, for MCDB and Misc, it almost looks like the very least skewed elements are way down at tiny abundances?

Do percentiles scan with S/N?

```{r avg abund}

all_di <- mutate(all_di, avg_abund = n0/s0)

skew_avgabund <- ggplot(data = all_di, aes(x = avg_abund, y = skew_percentile)) +
  geom_point(alpha = .3) +
  theme_bw() +
  ggtitle("Skew") +
  facet_wrap(vars(dat), scales = "free")

skew_avgabund

simpson_avgabund <- ggplot(data = all_di, aes(x = avg_abund, y = simpson_percentile)) +
  geom_point(alpha = .3) +
  theme_bw() +
  ggtitle("Simpson") +
  facet_wrap(vars(dat), scales = "free")

simpson_avgabund

```
There is perhaps a constraint, but that might be because you rarely get v high average abundances. You seem to rarely get low skewness or high Simpson at high average abundance, which scans with the heatmaps.

Do Simpson and skewness correspond?
```{r do skew and simp correlate}

ss_plot <- ggplot(data = all_di, aes(x= skew_percentile, y = simpson_percentile, color = dat)) +
  geom_point(alpha = .1) +
  scale_color_viridis_d(end = .8) +
  theme_bw() +
  ggtitle("Skew v Simp")

ss_plot

ss_facet_plot <- ss_plot +
  facet_wrap(vars(dat))

ss_facet_plot
```

There's a weak relationship between Simpson and skewness, but a lot of noise. They are not substitutable.

Why is Gentry so strangely bimodal? It's like a U, when all the others tend towards one end or the other.

Why are BBS and mammals less frequently squished than Portal?
